Haul-out behavior and aerial survey detectability of seals in the Bering and Chukchi seas

### 2022-01-13 analysis/paper/London_HauloutBeahvior.Rmd
WEATHER COVARIATE PLOTS/CAPTIONS FIXED

p <- ggplot(d,aes(day,hour,fill=percent_dry))+
  geom_tile(color= "white",size=0) + 
  scale_fill_distiller(palette = "YlGnBu",
                       name = "hourly percent dry", aesthetics = "fill",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                           barheight = 0.5, title.hjust = 0.5)
    )

@@

Visualizing the marginal or conditional effect of an individual weather
covariate on haul-out probability is difficult in this analysis because of the
collinearity between covariates as well as the spatial and temporal variation
across such a large region. The relationship of each weather covariate with
haul-out probability, averaged over the other weather conditions, is more
variable than simply showing model coefficients would imply. To elucidate
underlying patterns between the model's predicted haul-out probability and
specific weather covariates, we explored a subset of predictions during three
separate time windows. The same set of predictions used to create the haul-out
probability surfaces was reduced to include only local solar hours between 06:00
and 18:00 during three 7-day time windows. The time windows were centered on the
day of maximum haul-out probability and 25 days before and 25 days after peak
haul out. While this approach does not entirely remove patterns of collinearity
from the visualization (e.g. temperature increases with day of year), by
limiting our range of days and hours we gain some insight into how haul-out
probability varies across different weather covariate values. Figures
\@ref(fig:beardedHOwx), \@ref(fig:ribbonHOwx), and \@ref(fig:spottedHOwx)
present the predicted haul-out probability with respect to age and sex class
across the range of temperature and wind conditions encountered in the three
time windows. The visualizations also include vertical lines representing 95%
confidence intervals around the predicted haul-out probability to better
communicate the variation in model uncertainty. Additionally, simple linear fits
are shown to assist interpretation of overall patterns.

@@

Age and sex class were not included in the model for bearded seals due to our
lower sample size for adult and young-of-year age classes. As such,
results are shown for all ages. Bearded seals had a bi-modal
distribution of haul-out probability across the day (Figure
\@ref(fig:beardedHOCal)). In addition to a peak around local solar noon, the
bearded seal model predicts additional haul-out activity around local midnight.
The haul-out behavior is also more protracted throughout the spring season
compared to ribbon and spotted seals.

@@

```{r beardedHOwx,fig.height = 5.5, include=TRUE, fig.id = "beardedHOwx", fig.cap="Variability in predicted haul-out probability of bearded seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

@@

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Variability in predicted haul-out probability of ribbon seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

@@

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Variability in the predicted haul-out probability of spotted seals across the range of weather conditions encountered in the observed data. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

>>>>

p <- ggplot(d,aes(day,hour,fill=percent_dry))+
  geom_tile(color= "white",size=0) + 
  scale_fill_gradientn(colors = rev(met.brewer("Hiroshige")),
                       name = "hourly percent dry", aesthetics = "fill",
    guide = guide_colorbar(title.position = "bottom", barwidth = 15,
                           barheight = 0.5, title.hjust = 0.5)
    )

@@

Visualizing the marginal or conditional effect of an individual weather
covariate on haul-out probability is difficult in this analysis because of the
collinearity between covariates as well as the spatial and temporal variation
across such a large region. The relationship of each weather covariate with
haul-out probability, averaged over the other weather conditions, is more
variable than simply showing model coefficients would imply. That said,
important insights can be gained from plots of marginal effects. To create these
plots (figures \@ref(fig:beardedHOwx), \@ref(fig:ribbonHOwx), and
\@ref(fig:spottedHOwx)), we predicted haul-out probability across the full range
of weather covariate values while fixing hour of the day at local solar noon and
day-of-year at the average day of peak haul-out for each species. The
visualizations also include vertical lines representing 95% confidence intervals
around the predicted haul-out probability to better communicate the variation in
model uncertainty.

@@

Age and sex class were not included in the model for bearded seals due to our
lower sample size for adult and young-of-year age classes. As such, results are
shown for all ages. Overall, bearded seals were less likely to haul out and had
a bi-modal distribution of haul-out probability across the day (Figure
\@ref(fig:beardedHOCal)). In addition to a peak around local solar noon, the
bearded seal model predicts additional haul-out activity around local midnight.
In concert with the lower magnitude of haul-out probability, bearded seal
haul-out behavior is also more protracted throughout the spring season compared
to ribbon and spotted seals.

@@

```{r beardedHOwx,fig.height = 5.5, include=TRUE, fig.id = "beardedHOwx", fig.cap="Marginal effects of temperature, wind, pressure, and accumulated precipitation on the predicted haul-out probability of bearded seals combined across all age and sex classifications. Hour of the day was fixed at local solar noon and day-of-year held at the day of peak haul-out for bearded seals. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

@@

```{r ribbonHOwx, fig.height = 5.5, include=TRUE, fig.id = "ribbonHOwx", fig.cap="Marginal effects of temperature, wind, pressure, and accumulated precipitation on the predicted haul-out probability of ribbon seals within each age and sex classification. Hour of the day was fixed at local solar noon and day-of-year held at the average day of peak haul-out for ribbon seals. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}

@@

```{r spottedHOwx,fig.height = 5.5, include=TRUE, fig.id = "spottedHOwx", fig.cap="Marginal effects of temperature, wind, pressure, and accumulated precipitation on the predicted haul-out probability of spotted seals within each age and sex classification. Hour of the day was fixed at local solar noon and day-of-year held at the average day of peak haul-out for spotted seals. Transparent vertical lines represent the 95% confidence interval around the predicted haul-out probability."}
